A Compound Decision Approach to Covariance Matrix Estimation

### 2021-09-27 msg.tex

Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
\begin{equation}
\label{linear model}
\widehat{\bs{\Sigma}}_{\text{LW}} = \widehat{\rho}_{\text{LW}} \bs{S} + (1-\widehat{\rho}_{\text{LW}})\widehat{\mu} \bs{I}
\end{equation} 
where $\widehat{\mu} = \text{tr}(\bs{S})/p$, and $\widehat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\widehat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$. Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \eqref{frobenius risk}. We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
\begin{equation}
  \label{linear class}
  \beta_S s_{jk} + \beta_I u_{jk},
\end{equation}
where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. As proved in appendix, we can show that
$$URE(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
is an unbiased estimator of the risk \eqref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n(n-1)$ is an unbiased estimator of variance of $s_{jk}$. The optimal estimator in this class can now be chosen by minimizing $URE(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}. Proposition \eqref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \eqref{linear model}.
\begin{prop}
  \label{prop:linear}
  Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{R}(\beta_S, \beta_I)$. Then the coefficients $\widehat{\beta}_S$, $\widehat{\beta}_I $ and $\widehat{\rho}_{\text{LW}}$, $(1-\widehat{\rho}_{\text{LW}})\widehat{\mu}$ are very closed.

  \end{prop}

\subsection{\label{sec:proposed}Proposed estimator}

@@

Based on this result, we propose the following empirical Bayes procedure. We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the priors $G_{nd}$ and $G_d$. Under the Bayesian model \eqref{bayesian}, and the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent across $jk$, we estimate $G_{nd}$ using
\begin{equation}
  \label{Gnd hat}
  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
  \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG_d(a)\},
\end{equation}
where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $G_d$ is determined by $G_{nd}$ as indicated in \eqref{eq:marginal}. Of course, the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are not independent, so $\widehat{G}_{nd}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\widehat{G}_{nd}$ and $\widehat{G}_d$, we estimate the vectorized $\bs{\Sigma}$ using 
$$ \widehat{\bs{\delta}}(\bs{X})=
  (\widehat{t}(\bs{X}_{\cdot 2},\bs{X}_{\cdot 1}), \ldots,  \widehat{t}(\bs{X}_{\cdot p},\bs{X}_{\cdot (p-1)}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$$
  , where 
\begin{equation}
  \label{proposed}
  \widehat{t}(\bs{X}_{\cdot j},\bs{X}_{\cdot k}) = \frac{\int g(\bs{\eta}) f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
  \quad 
  \widehat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\widehat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\widehat{G}_d(a)}.
\end{equation}
The $\widehat{t}$ estimates the Bayes rule $t^\star$, $\widetilde{t}^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.

@@

\item Model 4. This setting was the same as Model 3 except with correlation parameter 0.9. This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gp hat}.

\item Model 5. With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4. This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.

\item Model 6. With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1. 
\end{itemize}

@@

\begin{table}
\begin{center}
\caption{\label{tab:sim1_time} Average running time for different ratios.}
%\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{lrrr}
\Hline
            & p=30 & p=100 & p=200 \\
\hline
Exemplar method   & 0.0748          & 4.2575         & 47.8616       \\
$K=2p$	      & 0.0445 	     & 1.3166	   & 12.2612         \\
$K=p$            & 0.0291         & 0.8744         & 8.0862         \\
$K=p/2$         & 0.0207         & 0.5861         & 5.3757         \\
$K=p/4$      & 0.0153         &0.4132          & 3.7415         \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \eqref{separable}. It equals our proposed estimator \eqref{proposed} except with the true $G_p$ \eqref{bayesian} instead of $\widehat{G}_p$ \eqref{Gp hat}. The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
\end{itemize}

@@

Figure \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators. Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9. This is not surprising because our method assumed independence of $\bs{A}_{jk}$.  In some cases, for example in Models 1 and 2, the improvement was substantial. Model 2 was especially interesting because the standard deviation and correlations were related. Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian} and leverage it to provide much more accurate estimates. The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5. In every setting, correcting MSG for positive-definiteness never increased the risk and decreased the risk in some cases. We also did experiments for Spectral norm and Matrix $\ell_1$. The results for this two norms are very similar to Frobenius norm. One exception is Adap has the lowest error in terms of Matrix $\ell_1$ norm in Model 1 because of its sparsity. Though our estimator was motivated in terms of the Frobenius norm error, it performed extremely well in terms of the other two norms as well.

@@

\begin{table}
\begin{center}
\caption{\label{tab:tab1} Average gene expression covariance matrix estimation errors. Bold entries highlight the smallest errors in each column.}
%\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{lrrr}
\Hline
            & Frobenius & Spectral & Matrix $\ell_1$ \\
\hline
MSG   & 24.07          & 12.89         & 47.12         \\
MSGCor & \textbf{23.90} & \textbf{12.87}& \textbf{46.63}         \\
Adap        & 28.81          & 17.20         & 50.71         \\
Linear      & 24.59         & 13.57         & 51.43         \\
QIS         & 28.66          & 15.94         & 59.30         \\
NERCOME     & 29.76          & 20.13         & 57.30         \\
Sample      & 28.39          & 16.10         & 57.90         \\
\hline

@@

Finally, we have so far assumed that our data are multivariate normal. To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gp hat} and in our proposed estimator \eqref{proposed}. If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.

@@

\begin{proof}
The estimator in \citet{ledoit2004well} is
\begin{equation} 
\label{eq:lw}
\bs{\Sigma}_{LW} = \frac{d_n^2-b_n^2}{d_n^2}\bs{S}+ \frac{b_n^2}{d_n^2}\hat{\mu}\bs{I}
\end{equation}
where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{n^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.

  We first rewrite the risk estimate $\widehat{R}(\beta_S, \beta_I)$. Define $\bs{M}=(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$. Then the unbiased risk estimator can be re-written as
  \[
 p^2 \widehat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
  \]
  where $\bs{Z} =(\bs{v_S}, \bs{v_I})$. Therefore
  \begin{align*}
    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\widehat{R}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
    &
    \widehat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\widehat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
  \end{align*}

@@

Therefore
\begin{align*}
\widehat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z^\top \bs{v_S}}-\bs{M} ) \\
&= \frac{1}{p^2d_n^2}\begin{pmatrix}
p & -\sum_{j=1}^p s_{jj}\\
-\sum_{j=1}^p s_{jj} & \sum_{j,k=1}^ps_{jk}^2
\end{pmatrix}\begin{pmatrix}
\sum_{j,k=1}^p(s_{jk}^2 - \widehat{\Delta}_{jk}^2)\\
\sum_{j=1}^p s_{jj}
\end{pmatrix}\\
&= \frac{1}{p^2d_n^2}\begin{pmatrix}
 p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
 (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)
\end{pmatrix}.
\end{align*}
The second component of $\widehat{\bs{\beta}}$ equals $\widehat{\beta}_I$, so
\begin{align*}
\widehat{\beta}_I &= \frac{1}{p^2d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)\\
&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p \} / d_n^2
=\widehat{\mu}\frac{b_n^2}{d_n^2}\frac{n}{n-1}.
\end{align*}
which is very closed to the coefficient of $\bs{I}$ in \eqref{eq:lw}
Furthermore,
\begin{align*}
\widehat{\beta}_I /\widehat{\mu} + \widehat{\beta}_S &= \frac{1}{p^2d_n^2}\sum_{j,k=1}^p \{ps_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2\}\\
&=\frac{1}{p^2d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
=
1.
\end{align*}
\end{proof}

\subsection{Proof of Proposition \eqref{prop:bayes risk}}
\begin{proof}
Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$
\begin{align*}
R(\beta_{S},\beta_{I}) &=\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} \{ (\beta_{S}s_{jk}+\beta_Iu_{jk}-\sigma_{jk})^2 \} \\
& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} [s_{jk}-\sigma_{jk} - \{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}]^2\\
& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E}[(s_{jk}-\sigma_{jk})^2+ \bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2-2(1-\beta_{S})\bb{E}s_{jk}(s_{jk}-\sigma_{jk})]\\
& =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\widehat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2]
\end{align*}
\end{proof}
\label{lastpage}
\end{document}

>>>>

Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
\begin{equation}
\label{linear model}
\widehat{\bs{\Sigma}}_{\text{LW}} = \widehat{\rho}_{\text{LW}} \bs{S} + (1-\widehat{\rho}_{\text{LW}})\widehat{\mu} \bs{I}
\end{equation} 
where $\widehat{\mu} = \text{tr}(\bs{S})/p$, and $\widehat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\widehat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$. Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \eqref{frobenius risk}. We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
\begin{equation}
  \label{linear class}
  \beta_S s_{jk} + \beta_I u_{jk},
\end{equation}
where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. As proved in appendix, we can show that
$$\textsc{RE}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
is an approximately unbiased estimator of the risk \eqref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ is an approximately unbiased estimator of variance of $s_{jk}$. The optimal estimator in this class can now be chosen by minimizing $RE(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}. Proposition \eqref{prop:linear} shows that this subspace shrinkage estimator is closely related to the \citet{ledoit2004well} estimator \eqref{linear model}.
\begin{prop}
\label{prop:linear}
Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{RE}(\beta_S, \beta_I)$. Then the estimator \eqref{linear model} can be expressed as 
$$\widehat{\bs{\Sigma}}_{\text{LW}} = (\widehat{\beta}_S)_+ \bs{S} + \min\{\widehat{\beta}_I, \hat{\mu}\} \bs{I}$$.
where $\hat{\mu} = \frac{1}{p}\sum_{j=1}^p s_{jj}$
\end{prop}
The first coefficient, which is constraint to be nonnegative, is an analog of nonnegative James-Stein estimator.
\subsection{\label{sec:proposed}Proposed estimator}

@@

Based on this result, we propose the following empirical Bayes procedure. We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the priors $G_{nd}$ and $G_d$. Under the Bayesian model \eqref{bayesian nondiagonal} and \eqref{bayesian diagonal}, and the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent across $jk$, we estimate $G_{nd}$ using
\begin{equation}
  \label{Gnd hat}
  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
  \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG_d(a)\},
\end{equation}
where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $G_d$ is determined by $G_{nd}$ as indicated in \eqref{eq:marginal}. Of course, the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are not independent, so $\widehat{G}_{nd}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\widehat{G}_{nd}$ and $\widehat{G}_d$, we estimate the vectorized $\bs{\Sigma}$ using 
$$ \widehat{\bs{\delta}}(\bs{X})=
  (\widehat{t}(\bs{X}_{\cdot 2},\bs{X}_{\cdot 1}), \ldots,  \widehat{t}(\bs{X}_{\cdot p},\bs{X}_{\cdot (p-1)}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$$
  , where 
\begin{equation}
  \label{proposed}
  \widehat{t}(\bs{X}_{\cdot j},\bs{X}_{\cdot k}) = \frac{\int g(\bs{\eta}) f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
  \quad 
  \widehat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\widehat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\widehat{G}_d(a)}.
\end{equation}
The $\widehat{t}$ estimates the Bayes rule $t^\star$, $\widetilde{t}^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.

@@

\item Model 4. This setting was the same as Model 3 except with correlation parameter 0.9. This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gnd hat}.

\item Model 5. With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4. This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.

\item Model 6. With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1. 
\end{itemize}

@@

\begin{table}
\begin{center}
\caption{\label{tab:sim1_time} Average running time for different ratios.}
%\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{lrrr}
\Hline
            & p=30 & p=100 & p=200 \\
\hline
Exemplar method   & 0.1357          & 6.0338        & 91.8942       \\
$K=2p$	      & 0.0539 	     & 1.1883	   & 10.7687         \\
$K=p$            & 0.0357         & 0.7852         & 7.0807         \\
$K=p/2$         & 0.0222        & 0.5231         & 4.6736         \\
$K=p/4$      & 0.0158         &0.3694          & 3.2749         \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\item OracMSG: It equals our proposed estimator \eqref{proposed} except sample grid points are cluster centers of true parameters $(\sigma_j,\sigma_k,r_{jk})$.
\end{itemize}

@@

Figure \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators. Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9. This is not surprising because our method assumed independence of $\bs{A}_{jk}$.  In some cases, for example in Models 1 and 2, the improvement was substantial. Model 2 was especially interesting because the standard deviation and correlations were related. Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian nondiagonal} and leverage it to provide much more accurate estimates. The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5. In every setting, correcting MSG for positive-definiteness never increased the risk and decreased the risk in some cases. We also did experiments for Spectral norm and Matrix $\ell_1$. The results for this two norms are very similar to Frobenius norm. One exception is Adap has the lowest error in terms of Matrix $\ell_1$ norm in Model 1 because of its sparsity. Though our estimator was motivated in terms of the Frobenius norm error, it performed extremely well in terms of the other two norms as well.

@@

\begin{table}
\begin{center}
\caption{\label{tab:tab1} Average gene expression covariance matrix estimation errors. Bold entries highlight the smallest errors in each column.}
%\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{lrrr}
\Hline
            & Frobenius & Spectral & Matrix $\ell_1$ \\
\hline
MSG   & 24.13 (2.54)         & 13.14(2.96)         & 46.53(8.49)         \\
MSGCor & \textbf{24.04(2.56)} & \textbf{13.12(2.94)}& \textbf{46.00(8.67)}         \\
Adap        & 28.93(2.03)          & 17.54(3.85)         & 49.66(8.47)         \\
Linear      & 24.49(2.53)         & 13.65(3.14)         & 49.46(9.04)         \\
QIS         & 28.76(2.79)          & 16.09(3.35)         & 58.73(9.74)         \\
NERCOME     & 24.52(2.52)          & 13.19(3.07)         & 48.27(9.67)         \\
Sample      & 28.43(2.54)          & 16.21(3.15)         & 56.08(9.89)         \\
\hline

@@

Finally, we have so far assumed that our data are multivariate normal. To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}. If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.

@@

\begin{proof}
The estimator in \citet{ledoit2004well} is
\begin{equation} 
\label{eq:lw}
\bs{\Sigma}_{LW} = (\frac{d_n^2-b_n^2}{d_n^2})_+\bs{S}+ \min\{1,\frac{b_n^2}{d_n^2}\}\hat{\mu}\bs{I}
\end{equation}
where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{pn^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.

  We first rewrite the risk estimate $\widehat{RE}(\beta_S, \beta_I)$. Define $\bs{M}=(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$. Then the unbiased risk estimator can be re-written as
  \[
 p^2 \widehat{RE}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
  \]
  where $\bs{Z} =(\bs{v_S}, \bs{v_I})$. Therefore
  \begin{align*}
    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\widehat{RE}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
    &
    \widehat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\widehat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
  \end{align*}

@@

Therefore
\begin{align*}
\widehat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z^\top \bs{v_S}}-\bs{M} ) \\
&= \frac{1}{p^2d_n^2}\begin{pmatrix}
p & -\sum_{j=1}^p s_{jj}\\
-\sum_{j=1}^p s_{jj} & \sum_{j,k=1}^ps_{jk}^2
\end{pmatrix}\begin{pmatrix}
\sum_{j,k=1}^p(s_{jk}^2 - \widehat{\Delta}_{jk}^2)\\
\sum_{j=1}^p s_{jj}
\end{pmatrix}\\
&= \frac{1}{p^2d_n^2}\begin{pmatrix}
 p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
 (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)
\end{pmatrix}.
\end{align*}
The second component of $\widehat{\bs{\beta}}$ equals $\widehat{\beta}_I$, so
\begin{align*}
\widehat{\beta}_I &= \frac{1}{p^2d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)\\
&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p \} / d_n^2
=\widehat{\mu}\frac{b_n^2}{d_n^2}.
\end{align*}
which is very closed to the coefficient of $\bs{I}$ in \eqref{eq:lw}
Furthermore,
\begin{align*}
\widehat{\beta}_I /\widehat{\mu} + \widehat{\beta}_S &= \frac{1}{p^2d_n^2}\sum_{j,k=1}^p \{ps_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2\}\\
&=\frac{1}{p^2d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
=
1.
\end{align*}
\end{proof}

\subsection{Proof of Proposition \eqref{prop:bayes risk}}
\begin{proof}
Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$
\begin{align*}
R(\beta_{S},\beta_{I}) &=\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} \{ (\beta_{S}s_{jk}+\beta_Iu_{jk}-\sigma_{jk})^2 \} \\
& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} [s_{jk}-\sigma_{jk} - \{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}]^2\\
& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E}[(s_{jk}-\sigma_{jk})^2+ \bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2-2(1-\beta_{S})\bb{E}s_{jk}(s_{jk}-\sigma_{jk})]\\
& =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\frac{n}{n-1}\widehat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2]
\end{align*}
The last equality is based on the fact that $\text{Var}(s_{jk}) = \frac{1}{n(n-1)}\sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2$.
\end{proof}
\label{lastpage}
\end{document}
